The increasing cost of happiness

### 2021-06-07 src/draft-SSM1.Rmd
Discussion of Twenge added (closes issue #16)

Other researchers have argued income has a satiety point on affective well-being, after which increasing income has no further effect [@kahneman2010high; @jebb2018happiness]. This is consistent with a change point after which the slope is zero. While we observed a change-point, the slope afterwards was small but credibly larger than zero so our results are not strictly consistent with a "satiety" point. Our results emphasize the slope over income is not constant. Other researchers have argued that affective wellbeing is a constant (albiet decreasing) log-linear function of income [@killingsworth2021experienced; @sacks2012newl] or a constant linear relationship [@twenge2020expanding], however these studies typically used a binned measure of income which may reduce the sensitivity to change-points in the final functional form.

>>>>

Other researchers have argued income has a satiety point on affective well-being, after which increasing income has no further effect [@kahneman2010high; @jebb2018happiness]. This is consistent with a change point after which the slope is zero. While we observed a change-point, the slope afterwards was small but credibly larger than zero so our results are not strictly consistent with a "satiety" point. Our results emphasize the slope over income is not constant. Other researchers have argued that affective wellbeing is a constant (albiet decreasing) log-linear function of income [@killingsworth2021experienced; @sacks2012newl] or a constant linear relationship [@twenge2020expanding]. We also show the change-point has increased over time. Indeed Twenge et al report the linear relationship between income and happiness has become steeper over time in the US General Social Survey (GSS). While we did not see any consistent trend over time in pre or post-slope changes in the piecewise-linear model of affective-wellbeing (e.g., Figure 2), there was a credibly increasing slope in the linear model of cognitive-wellbeing with time (Table S3). It may be the single-item on general happiness in the GSS has more in common with the single item on life-satisfaction in HILDA than the more comprehensive 9-item, dual-component questions on affective-wellbeing in the SF-36. Nevertheless, like Twenge et al, we report that the difference in happiness betweeen the rich and poor has been increasing over time (e.g., Figure S6), however our piecewise-model reveals the income level around which this difference occurs.
